{\steve}: A hierarchical Bayesian model for Supernova Cosmology

### 2018-12-29 papers/methods/figures/tikz.tex
Updating paper based on referee comments

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{tikz}
\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
\usetikzlibrary{bayesnet}
\usetikzlibrary{external}
\tikzexternalize % activate!
\tikzset{external/force remake}
\begin{document}
	\begin{tikzpicture}
	% Obs nodes
	\node[obs]  (obssum) {$\hat{m}_B, \hat{x}_1, \hat{c}, \mathcal{C_\eta}^{-1}$};
	\node[obs, left=0.8cm of obssum]  (zobs) {$\hat{z}$};
	\node[obs, right=0.6cm of obssum]  (mobs) {$\hat{m}$};
	\node[obs, right=0.6cm of mobs]  (pobs) {$\hat{p}$};
	% Latent nodes
	\node[latent, above=of obssum] (sum) {$m_B, x_1, c$};
	\node[det, above=of zobs] (z) {$z$};	
	\node[det, above=of z] (mu) {$\mu$};
	\node[det, right=4.33cm of mu] (MBobs) {$M_B, x_1^*, c^*$};
	% Survey nodes
	\node[latent, above=of sum, yshift=1.6cm, xshift=1cm, text width=1.9cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\sigma_{M_B}, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c$ };
	\node[latent, right=1.2cm of pop1, align=center] (sf) {$\delta S$};
	\node[latent, left=0.8cm of pop1, align=center] (kappa) {$\kappa_0, \kappa_1$};
	% Global nodes
	\node[latent, above=2.3cm of mu]  (cosmo) {$\Omega_m, w$};
	\node[latent, above=1.7cm of MBobs, text width=2.5cm, align=center]  (ab) {$\langle M_B \rangle, \alpha, \beta$, $\delta(0)$, \\$\delta(\infty)$};
	\node[latent, right=0.5cm of cosmo, align=center]  (dz) {$\delta \mathcal{Z}_i$};
	% Connect the nodes
	\edge {sum} {obssum};
	\edge {z} {zobs};
	\edge {z,cosmo} {mu};
	\edge {mu, pop1, ab, sf} {MBobs};	
	\edge {MBobs} {pobs, mobs};
	\edge {dz, kappa, MBobs} {sum};
	% Plates
	\plated[thick] {sn} {(obssum)(zobs)(mobs)(pobs)(sum)(z)(mu)(MBobs)} {$\forall\ \rm{SN}$} ;
	\plated[thick] {survey} {(sn)(pop1)(kappa)(sf)} {\\$\forall\ \rm{Survey}$} ;
	\end{tikzpicture}
\end{document}

>>>>

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{tikz}
\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
\usetikzlibrary{bayesnet}
\usetikzlibrary{external}
\tikzexternalize % activate!
\tikzset{external/force remake}
\begin{document}
	\begin{tikzpicture}
	% Obs nodes
	\node[obs]  (obssum) {$\hat{m}_B, \hat{x}_1, \hat{c}, \mathcal{C_\eta}^{-1}$};
	\node[obs, left=0.8cm of obssum]  (zobs) {$\hat{z}$};
	\node[obs, right=0.6cm of obssum]  (mobs) {$\hat{m}$};
	\node[obs, right=0.6cm of mobs]  (pobs) {$\hat{p}$};
	% Latent nodes
	\node[latent, above=of obssum] (sum) {$m_B, x_1, c$};
	\node[det, above=of zobs] (z) {$z$};	
	\node[det, above=of z] (mu) {$\mu$};
	\node[latent, right=4.1cm of mu] (MBobs) {$M_B, x_1^*, c^*$};
	% Survey nodes
	\node[latent, above=of sum, yshift=1.6cm, xshift=1cm, text width=1.9cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\sigma_{M_B}, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c$ };
	\node[latent, right=1.2cm of pop1, align=center] (sf) {$\delta S$};
	\node[latent, left=0.8cm of pop1, align=center] (kappa) {$\kappa_0, \kappa_1$};
	% Global nodes
	\node[latent, above=2.3cm of mu]  (cosmo) {$\Omega_m, w$};
	\node[latent, above=2.3cm of MBobs, text width=2cm, align=center]  (ab) {$\langle M_B \rangle, \alpha, \beta$, \\$\delta(0)$, $\delta(\infty)$};
	\node[latent, right=0.7cm of cosmo, align=center]  (dz) {$\delta \mathcal{Z}_i$};
	% Connect the nodes
	\edge {sum} {obssum};
	\edge {z} {zobs};
	\edge {z,cosmo} {mu};
	\edge {mu, pop1, ab, sf} {MBobs};	
	\edge {MBobs} {pobs, mobs};
	\edge {dz, kappa, MBobs} {sum};
	% Plates
	\plated[thick] {sn} {(obssum)(zobs)(mobs)(pobs)(sum)(z)(mu)(MBobs)} {$\forall\ \rm{SN}$} ;
	\plated[thick] {survey} {(sn)(pop1)(kappa)(sf)} {\\$\forall\ \rm{Survey}$} ;
	\end{tikzpicture}
\end{document}

### 2018-12-29 papers/methods/paper.tex
Updating paper based on referee comments

As we are focused on the spectroscopically confirmed supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$ such that $P(\hat{z}|z) = \delta(\hat{z} - z)$. Potential sources of redshift error (such as peculiar velocities) are taken into account not via uncertainty on redshift (which is technically challenging to implement \added{as varying redshifts introduce computational complexity in computing the distance modulus integral}) but instead uncertainty on distance modulus. Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable to represent uncertainty on the probability estimate. One of the strengths of {\steve} (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, these misclassifications can naturally be modelled and taken into account by introducing additional populations that supernovae have a non-zero probability of belonging to.

@@

This method of creating a secondary covariance matrix using partial derivatives is used throughout the traditional and BHM analyses. For each survey and band, we have two systematics --- the calibration uncertainty and the filter wavelength uncertainty. We include these in our approach, in addition to including HST Calspec calibration uncertainty, ten SALT2 model systematic uncertainties, a dust systematic, a global redshift bias systematic, and also the systematic peculiar velocity uncertainty. A comprehensive explanation of all systematics is given in \citet{Brout18SYS}; see Table~4 for details. This gives thirteen global systematics shared by \added{ all surveys that apply globally to all supernova summary statistics}, plus two systematics per band in each survey. With $\eta \equiv \lbrace m_B, x_1, c \rbrace$, our initial conditional likelihood for our observed summary statistics shown in Equation \eqref{eq:pop} becomes
\begin{equation}
P\left(\hat{\eta}, \frac{\partial \hat{\eta}}{\partial \Z_i} | \eta, \delta \Z_i, C_\eta\right) = \mathcal{N}\left(\hat{\eta} + \delta \Z_i \frac{\partial \hat{\eta}}{\partial \Z_i}|\eta,\cov\right). \label{eq:l3}
\end{equation}

@@

The selection functions parameters (a skew normal for low-redshift and a complimentary error function for high-redshift) are all given independent uncertainty of $0.01$ (mean and width for the CDF selection function, and mean, width and skewness for the skew normal selection function). Draw from each survey simulation until we have 1000 low-$z$ supernovae and 1000 DES-like supernovae, representing a statistical sample of greater power than the estimated 350 supernovae for the DES-SN3YR sample. Sample data for 1000 high and low redshift supernovae are shown in Figure \ref{fig:simple_pop}, confirming the presence of strong selection effects in both toy surveys, as designed.

@@

\added{ The inclusion of a proper treatment of intrinsic scatter is significant, as} prior analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population \citep{Conley2011, Betoule2014}, but recent analyses require a more sophisticated approach. In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models via simulations, the {\gten} and {\celeven} models described in Section \ref{sec:challenges}. The {\gten} models dispersion with 70\% contribution from coherent variation and 30\% from chromatic variation whilst the {\celeven} model has 25\% coherent scatter and 75\% from chromatic variation. These two broadband scatter models are converted to spectral energy distribution models for use in simulations in \citet{Kessler13}.

@@

Difficulty in providing an adequate parametrisation for realistic intrinsic dispersion, and the simplification of Malmquist bias to only apparent magnitude also leads to biases in the population parameters. \added{ To determine if the underlying population mischaracterisation was a cause of cosmological bias, we ran fits wherein the underlying population was fixed to the known distributions used for the simulation. These fits did not show reduced cosmological bias in the {\celeven} simulations, and thus the determination of underlying population in our model may give biased distributions, however does not impact cosmology.} As the population parameters recovered using the simplistic toy supernova data in the previous section do not exhibit significant bias, future work will focus on intrinsic dispersion and Malmquist bias rather than alternate parametrisations of the underlying supernova population.

@@

\subsection{General Selection Effects}
\label{app:selection1}
When formulating and fitting a model using a constraining dataset, we wish to resolve the posterior surface defined by
\begin{equation}
P(\theta | {\rm data}) \propto P({\rm data} | \theta) P(\theta),
\end{equation}
which gives the probability of the model parameter values ($\theta$) given the data.  Prior knowledge of the allowed values of the model parameters is encapsulated in the prior probability $P(\theta)$. Of primary interest to us is the likelihood of observing the data given our parametrised model, $\mathcal{L} \equiv P({\rm data} | \theta)$. When dealing with experiments that have imperfect selection efficiency, our likelihood must take that efficiency into account.  We need to describe the probability that the events we observe are both drawn from the distribution predicted by the underlying theoretical model \textit{and} that those events, given they happened, are subsequently successfully observed.  To make this extra conditional explicit, we write the likelihood of the data given an underlying model, $\theta$, \textit{and} that the data are included in our sample, denoted by $S$, as:
\begin{equation}
\mathcal{L}(\theta; {\rm data}) = P({\rm data} | \theta, S). \label{eq:app_like}
\end{equation}
A variety of selection criteria are possible, and in our method we use our data in combination with the proposed model to determine the probability of particular selection criteria.  That is, we characterise a function $P(S|{\rm data},\theta)$, which colloquially can be stated as \textit{the probability of a potential observation passing selection cuts, given our observations and the underlying model}. We can introduce this expression in a few lines due to symmetry of joint probabilities and utilising that $P(x,y,z) = P(x|y,z)P(y,z) = P(y|x, z)P(x, z)$:
\begin{eqnarray}
P({\rm data} | S , \theta) P(S, \theta) &=& P(S | {\rm data}, \theta) P({\rm data}, \theta)\\
P({\rm data} | S , \theta) &=& \frac{ P(S | {\rm data}, \theta) P({\rm data}, \theta) }{ P(S, \theta) }\\
&=& \frac{ P(S | {\rm data}, \theta) P({\rm data} | \theta) P(\theta)}{ P(S | \theta)  P(\theta)}\\
&=& \frac{ P(S | {\rm data}, \theta) P({\rm data} | \theta)}{ P(S | \theta) } \label{eq:general1}
\end{eqnarray}
which is equal to the likelihood $\mathcal{L}$. \added{ At this point, our derivation depends on whether our selection effects are best modelled as a function of observables or as a function of latent parameters. In most cases, selection effects will be best modelled directly as a function of observables, and we would introducing an integral over all possible events $D$, so we can evaluate $P(S|\theta)$, 
\begin{eqnarray}
\mathcal{L}(\theta; {\rm data})&=& \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S, D|\theta)\, dD} \\
\mathcal{L}(\theta; {\rm data}) &=& \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}, \label{eq:app_main}
\end{eqnarray}
where we define the denominator as $d$ for simplicity in future derivations. The second option, wherein selection effects can be more efficiently modelled with the inclusion of latent parameters (for example, in the case where we do not have direct access to the observables upon which our data selection is determined), we can introduce our latent parameters in addition to an integral over all possible data:
\begin{eqnarray}
\mathcal{L}(\theta; {\rm data})&=& \frac{\int P(S| L,\theta) P({\rm data}|L) P(L|\theta)\, dL}{\iint P(S|D,L,\theta)P(D|L,\theta) P(L|\theta)\, dD\, dL},\label{eq:latent}
\end{eqnarray}
where $L$ represents our latent parameters that model the true underlying values of our observables, such that our data is conditioned directly on them.
}

\subsection{Supernova Selection Effects}
\label{app:selection2}
\added{
To turn the generalised equations from the previous sections into selection effects relevant for this model, need to highlight that $\theta$ represents only our top level parameters ($\Omega_m$, $w$, $\alpha$, $\beta$, etc), and that our parametrisation of true underlying values (for example $m_B$) takes the form of latent parameters. We thus continue from equation \eqref{eq:latent} and write out denominator
\begin{equation}
d = \iint  P(S|D, L, \theta) P(D | L, \theta) P(L | \theta)\, dD\, dL
\end{equation}
Here we can now move from the generic labels $D$ and $L$ to something specific to our model. In this case, $D$ would represent our observables, $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{m}, \hat{p}\rbrace$, and $L$ our latent parameters $\lbrace m_B, x_1, c, z\rbrace.$ However, as we assume that our selection effects can be sufficiently encapsulated by latent parameters. That is, we simplify $P(S|D,L, \theta) \rightarrow P(S|L)$. Specifically, we assume that our selection effects can be quantified using the true apparent magnitude $m_B$ and redshift $z$, and due to this we will not write out the other parameters in our model for simplicity. If we did, we'd simply find them integrated to unity. Formally, we assume that $P(S|D, L, \theta) = P(S|z) P(S|m_B)$.}
Writing out the relevant observables $\hat{m}_B$, $\hat{z}$ and latent parameters $m_B$ and $z$, our denominator then becomes
\begin{equation}
d = \int P(S|z) P(S|m_B) P(\hat{z}|z) P(\hat{m}_B|m_B) P(z, m_B | \theta)\, d\hat{z} \, d\hat{m}_B \, dz \, dm_B.  \label{eq:sel} 
\end{equation}
Due to our assumed perfect measurement of redshift, $P(\hat{z}|z) = \delta(\hat{z} - z)$. $P(\hat{m}_B | m_B)$ is a Gaussian due to our Gaussian model of summary statistics, and can be analytically integrated out, collapsing the integral over $\hat{m}_B$. Finally, we can express $P(z, m_B | \theta)$ as  $P(m_B | z, \theta) P(z | \theta)$, where the first term requires us to calculate the magnitude distribution of our underlying population at a given redshift, and the second term is dependent on survey geometry and supernovae rates. We can thus state
\begin{equation}
d = \int \left[ \int P(S|m_B) P(m_B | z, \theta)\, dm_B \right] P(S|z)P(z|\theta)\, dz.
\end{equation}
By assuming that the distribution $P(S|z)P(z|\theta)$ is well sampled by the observed supernova redshifts, we can approximate the integral over redshift by evaluating
\begin{equation}
\int P(S|m_B) P(m_B | z, \theta)\, dm_B \label{eq:selint}
\end{equation}
for each supernova in the dataset -- i.e. Monte Carlo integration with assumed perfect importance sampling.

>>>>

As we are focused on the spectroscopically confirmed supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$ such that $P(\hat{z}|z) = \delta(\hat{z} - z)$. Potential sources of redshift error (such as peculiar velocities) are taken into account not via uncertainty on redshift (which is technically challenging to implement \added{as varying redshifts introduce computational complexity in computing the distance modulus integral by reducing the amount of pre-computation that can be utilised}) but instead uncertainty on distance modulus. Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable to represent uncertainty on the probability estimate. One of the strengths of {\steve} (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, these misclassifications can naturally be modelled and taken into account by introducing additional populations that supernovae have a non-zero probability of belonging to.

@@

This method of creating a secondary covariance matrix using partial derivatives is used throughout the traditional and BHM analyses. For each survey and band, we have two systematics --- the calibration uncertainty and the filter wavelength uncertainty. We include these in our approach, in addition to including HST Calspec calibration uncertainty, ten SALT2 model systematic uncertainties, a dust systematic, a global redshift bias systematic, and also the systematic peculiar velocity uncertainty. A comprehensive explanation of all systematics is given in \citet{Brout18SYS}; see Table~4 for details. This gives thirteen global systematics shared by \added{ all surveys that apply globally to all supernova summary statistics}, plus two systematics per band in each survey. \added{Systematics with known correlations are shifted together to produce covariant deviations, and we thus assume that the numerical derivatives input into our model represent independent systematics. Full details can be found in \citet{Brout18SYS}.} With $\eta \equiv \lbrace m_B, x_1, c \rbrace$, our initial conditional likelihood for our observed summary statistics shown in Equation \eqref{eq:pop} becomes
\begin{equation}
P\left(\hat{\eta}, \frac{\partial \hat{\eta}}{\partial \Z_i} | \eta, \delta \Z_i, C_\eta\right) = \mathcal{N}\left(\hat{\eta} + \delta \Z_i \frac{\partial \hat{\eta}}{\partial \Z_i}|\eta,\cov\right). \label{eq:l3}
\end{equation}

@@

The selection functions parameters (a skew normal for low-redshift and a \replaced{complimentary}{complementary} error function for high-redshift) are all given independent uncertainty of $0.01$ (mean and width for the CDF selection function, and mean, width and skewness for the skew normal selection function). Draw from each survey simulation until we have 1000 low-$z$ supernovae and 1000 DES-like supernovae, representing a statistical sample of greater power than the estimated 350 supernovae for the DES-SN3YR sample. Sample data for 1000 high and low redshift supernovae are shown in Figure \ref{fig:simple_pop}, confirming the presence of strong selection effects in both toy surveys, as designed.

@@

Prior analyses often treated intrinsic dispersion simply as scatter in the underlying absolute magnitude of the underlying population \citep{Conley2011, Betoule2014}, but recent analyses require a more sophisticated approach. In our development of this model and tests of intrinsic dispersion, we analyse the effects of two different scatter models via simulations, the {\gten} and {\celeven} models described in Section \ref{sec:challenges}. The {\gten} models dispersion with 70\% contribution from coherent variation and 30\% from chromatic variation whilst the {\celeven} model has 25\% coherent scatter and 75\% from chromatic variation. These two broadband scatter models are converted to spectral energy distribution models for use in simulations in \citet{Kessler13}.

@@

Difficulty in providing an adequate parametrisation for realistic intrinsic dispersion, and the simplification of Malmquist bias to only apparent magnitude also leads to biases in the population parameters. \added{ To determine if the underlying population mischaracterisation was a cause of cosmological bias, we ran fits wherein the underlying population was fixed to the known distributions used for the simulation. These fits did not change the bias in the {\celeven} simulation. We conclude that the biased population recovery is not the cause of cosmological bias.} As the population parameters recovered using the simplistic toy supernova data in the previous section do not exhibit significant bias, future work will focus on intrinsic dispersion and Malmquist bias rather than alternate parametrisations of the underlying supernova population.

@@

\subsection{General Selection Effects}
\label{app:selection1}
When formulating and fitting a model using a constraining dataset, we wish to resolve the posterior surface defined by
\begin{equation}
P(\theta | {\rm data}) \propto P({\rm data} | \theta) P(\theta),
\end{equation}
which gives the probability of the model parameter values ($\theta$) given the data.  Prior knowledge of the allowed values of the model parameters is encapsulated in the prior probability $P(\theta)$. Of primary interest to us is the likelihood of observing the data given our parametrised model, $\mathcal{L} \equiv P({\rm data} | \theta)$. When dealing with experiments that have imperfect selection efficiency, our likelihood must take that efficiency into account.  We need to describe the probability that the events we observe are both drawn from the distribution predicted by the underlying theoretical model \textit{and} that those events, given they happened, are subsequently successfully observed.  To make this extra conditional explicit, we write the likelihood of the data given an underlying model, $\theta$, \textit{and} that the data are included in our sample, denoted by $S$, as:
\begin{equation}
\mathcal{L}(\theta; {\rm data}) = P({\rm data} | \theta, S). \label{eq:app_like}
\end{equation}
A variety of selection criteria are possible, and in our method we use our data in combination with the proposed model to determine the probability of particular selection criteria.  That is, we characterise a function $P(S|{\rm data},\theta)$, which colloquially can be stated as \textit{the probability of a potential observation passing selection cuts, given our observations and the underlying model}. We can introduce this expression in a few lines due to symmetry of joint probabilities and utilising that $P(x,y,z) = P(x|y,z)P(y,z) = P(y|x, z)P(x, z)$:
\begin{eqnarray}
P({\rm data} | S , \theta) P(S, \theta) &=& P(S | {\rm data}, \theta) P({\rm data}, \theta)\\
P({\rm data} | S , \theta) &=& \frac{ P(S | {\rm data}, \theta) P({\rm data}, \theta) }{ P(S, \theta) }\\
&=& \frac{ P(S | {\rm data}, \theta) P({\rm data} | \theta) P(\theta)}{ P(S | \theta)  P(\theta)}\\
&=& \frac{ P(S | {\rm data}, \theta) P({\rm data} | \theta)}{ P(S | \theta) } \label{eq:general1}
\end{eqnarray}
which is equal to the likelihood $\mathcal{L}$. \added{ At this point, our derivation depends on whether our selection effects are best modelled as a function of observables or as a function of latent parameters. In most cases, selection effects will be best modelled directly as a function of observables, and we would introducing an integral over all possible events $D$, so we can evaluate $P(S|\theta)$, 
\begin{eqnarray}
\mathcal{L}(\theta; {\rm data})&=& \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S, D|\theta)\, dD} \\
\mathcal{L}(\theta; {\rm data}) &=& \frac{P(S|{\rm data},\theta) P({\rm data}|\theta)}{\int P(S | D, \theta) P(D|\theta)\, dD}. \label{eq:app_main}
\end{eqnarray}
The second option, wherein selection effects can be more efficiently modelled with the inclusion of latent parameters (for example, in the case where we do not have direct access to the observables upon which our data selection is determined), we can introduce our latent parameters in addition to an integral over all possible data:
\begin{eqnarray}
\mathcal{L}(\theta; {\rm data})&=& \frac{\int P(S| L,\theta) P({\rm data}|L) P(L|\theta)\, dL}{\iint P(S|D,L,\theta)P(D|L,\theta) P(L|\theta)\, dD\, dL},\label{eq:latent}
\end{eqnarray}
where $L$ represents our latent parameters that model the true underlying values of our observables, such that our data is conditioned directly on them.
}

\subsection{Supernova Selection Effects}
\label{app:selection2}
\added{
To turn the generalised equations from the previous sections into selection effects relevant for this model, need to highlight that $\theta$ represents only our top level parameters ($\Omega_m$, $w$, $\alpha$, $\beta$, etc), and that our parametrisation of true underlying values (for example $m_B$) takes the form of latent parameters. We thus continue from equation \eqref{eq:latent} and write out denominator $d$:
\begin{equation}
d = \iint  P(S|D, L, \theta) P(D | L, \theta) P(L | \theta)\, dD\, dL.
\end{equation}
Here we can now move from the generic labels $D$ and $L$ to something specific to our model. In this case, $D$ would represent our observables, $\lbrace \hat{m}_B, \hat{x}_1, \hat{c}, \hat{z}, \hat{m}, \hat{p}\rbrace$, and $L$ our latent parameters $\lbrace m_B, x_1, c, z\rbrace.$ We assume that our selection effects can be sufficiently encapsulated by latent parameters. That is, we simplify $P(S|D,L, \theta) \rightarrow P(S|L)$. Specifically, we assume that our selection effects can be quantified using the true apparent magnitude $m_B$ and redshift $z$, and due to this we will not write out the other parameters in our model for simplicity. If we did, we would simply find them integrated to unity. Formally, we assume that $P(S|D, L, \theta) = P(S|z) P(S|m_B)$.}
Writing out the relevant observables $\hat{m}_B$, $\hat{z}$ and latent parameters $m_B$ and $z$, our denominator then becomes
\begin{equation}
d = \int P(S|z) P(S|m_B) P(\hat{z}|z) P(\hat{m}_B|m_B) P(z, m_B | \theta)\, d\hat{z} \, d\hat{m}_B \, dz \, dm_B.  \label{eq:sel} 
\end{equation}
Due to our assumed perfect measurement of redshift, $P(\hat{z}|z) = \delta(\hat{z} - z)$. $P(\hat{m}_B | m_B)$ is a Gaussian due to our Gaussian model of summary statistics, and can be analytically integrated out, collapsing the integral over $\hat{m}_B$. Finally, we can express $P(z, m_B | \theta)$ as  $P(m_B | z, \theta) P(z | \theta)$, where the first term requires us to calculate the magnitude distribution of our underlying population at a given redshift, and the second term is dependent on survey geometry and supernovae rates. We can thus state
\begin{equation}
d = \int \left[ \int P(S|m_B) P(m_B | z, \theta)\, dm_B \right] P(S|z)P(z|\theta)\, dz.
\end{equation}
By assuming that the distribution $P(S|z)P(z|\theta)$ is well sampled by the observed supernova redshifts, we can approximate the integral over redshift by evaluating
\begin{equation}
\int P(S|m_B) P(m_B | z, \theta)\, dm_B \label{eq:selint}
\end{equation}
for each supernova in the dataset -- i.e. Monte Carlo integration with assumed perfect importance sampling.
